---
title: "qiime2R_Anoles"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r, echo = T, eval = T}

#BiocManager::install("dada2")

#load packages

library(dada2)
library(phyloseq)
library(Biostrings)
library(ggplot2)
library(dbplyr)
library(tidyverse)
library(qiime2R)
library(plyr)
library(ANCOMBC)
library(microbiome)
#library(nplotr)
library(rmeta)
library(readr)
library(vegan)
```

## Make functions to convert Phyloseq objects into vegan objects

``` {r, echo = T, eval = T}
# convert the sample_data() within a phyloseq object into a vegan compatible data object
pssd2veg <- function(physeq) {
  sd <- sample_data(physeq)
  return(as(sd,"data.frame"))
}

#convert the asv table (otu_table) within a phyloseq object into a vegan compatible data object
psotu2veg <- function(physeq) {
  OTU <- otu_table(physeq)
  if (taxa_are_rows(OTU)) {
    OTU <- t(OTU)
  }
  return(as(OTU, "matrix"))
}



options(max.print = 5000)
```


## qiime 2 R import and conversion into phyloseq object
``` {r, echo = T, eval = T}

#import qiime2 data into a phyloseq object
allps<-qza_to_phyloseq(features="all_feature-table.qza", taxonomy = "seqs-taxonomy.qza", metadata="AHC_meta.tsv")

# convert sequences to ASV names and create new field to keep the reference sequences
dna <- Biostrings::DNAStringSet(taxa_names(allps))
names(dna) <- taxa_names(allps)
allps <- merge_phyloseq(allps, dna)
taxa_names(allps) <- paste0("ASV", seq(ntaxa(allps)))

sample_data(allps)

```



## breakdown the full ps object then prune samples

``` {r, echo = T, eval = T}
# ID samples to prune from anoles = AF107, AF008, and AF086
psM <- subset_samples(allps, env_feature == "AnoleFeces") %>% subset_samples(scientific_name == "cristatellus") %>% subset_samples(description == "Mayaguez Forest" | description == "Mayaguez Urban" | description =="Maricao Rocks" | description =="AirBnB")
taxa_barpsM = plot_bar(psM, fill = "Phylum", x = "ID")
psM.prune <- subset_samples(psM, ID !="AF107" & ID !="AF008" & ID !="AF086")


# ID samples to prune from cyotes = 42F, 18F, 53F. Kept the specific ones for now and pruned sub5k for now
psC <- subset_samples(allps, env_feature == "CoyoteFeces")
taxa_barpsC = plot_bar(psC, fill = "Phylum", x = "ID")
taxa_barpsC + theme(legend.position = "none")
psC.prune <- prune_samples(sample_sums(psC)>=5000, psC)

#ID samples to prune from humans and filter out children.
psH <- subset_samples(allps, env_feature == "HumanFeces") %>% subset_samples(life_stage != "child")
taxa_barpsH = plot_bar(psH, fill = "Phylum", x = "ID")
taxa_barpsH + theme(legend.position = "none")
psH.prune <- prune_samples(sample_sums(psH)>=300000, psH)







```


## Make all of the required ps objects. First combine and then rarefy

``` {r, echo = T, eval = T}
#Maya anoles
psM.rare <- rarefy_even_depth(psM.prune, rngseed=1, sample.size=0.9*min(sample_sums(psM.prune)), replace=F)
plot_bar(psM.rare, fill = "Phylum", x = "ID")
#adult USA Human and Maya anoles 
psUM <- subset_samples(psH.prune, description =="USA") %>% merge_phyloseq(psM.prune)
sample_data(psUM)
psUM.rare = rarefy_even_depth(psUM, rngseed=1, sample.size=0.9*min(sample_sums(psUM)), replace=F)


#adult Venezuela Human and Maya anoles 
psVM <- subset_samples(psH.prune, description =="Venezuela") %>% merge_phyloseq(psM.prune)
sample_data(psVM)
psVM.rare = rarefy_even_depth(psVM, rngseed=1, sample.size=0.9*min(sample_sums(psVM)), replace=F)


#adult Malawi Human and Maya anoles 
psMAM <- subset_samples(psH.prune, description =="Malawi") %>% merge_phyloseq(psM.prune)
psMAM.rare = rarefy_even_depth(psMAM, rngseed=1, sample.size=0.9*min(sample_sums(psMAM)), replace=F)


#adult USA Human, Maya anoles, and coyotes with duodenal
psUMC.duo <- subset_samples(psH.prune, description =="USA") %>% merge_phyloseq(psM.prune, psC.prune)
psUMC.duo.rare = rarefy_even_depth(psUMC.duo, rngseed=1, sample.size=0.9*min(sample_sums(psUMC.duo)), replace=F)


#adult USA Human, Maya anoles, and coyotes without duodenal
psUMC <- subset_samples(psUMC.duo, description != "Urban coyote duodenal samples")
psUMC.rare = rarefy_even_depth(psUMC, rngseed=1, sample.size=0.9*min(sample_sums(psUMC)), replace=F)

#coyotes
psC.rare <- rarefy_even_depth(psC.prune, rngseed=1, sample.size=0.9*min(sample_sums(psC.prune)), replace=F)

#Everything other than duodenal and children
psAll.prune <- merge_phyloseq(psM.prune, psC.prune, psH.prune) %>% subset_samples(description != "Urban coyote duodenal samples")
psAll.rare = rarefy_even_depth(psAll.prune, rngseed=1, sample.size=0.9*min(sample_sums(psAll.prune)), replace=F)


#adult Venezuela Human, Maya anoles, and non duo coyptes
psVMC <- merge_phyloseq(psVM, psC.prune) %>% subset_samples(description != "Urban coyote duodenal samples")
psVMC.rare = rarefy_even_depth(psVMC, rngseed=1, sample.size=0.9*min(sample_sums(psVMC)), replace=F)


#adult Malawi Human, Maya anoles, and non duo coyptes
psMAMC <- merge_phyloseq(psMAM, psC.prune) %>% subset_samples(description != "Urban coyote duodenal samples")
psMAMC.rare = rarefy_even_depth(psMAMC, rngseed=1, sample.size=0.9*min(sample_sums(psMAMC)), replace=F)


#adult USA and Malawi Humans
psANCOM.USA.Malawi <- subset_samples(psH.prune, description =="Malawi" | description =="USA")
psANCOM.USA.Malawi.rare <- rarefy_even_depth(psANCOM.USA.Malawi, rngseed=1, sample.size=0.9*min(sample_sums(psANCOM.USA.Malawi)), replace=F)

#adult USA and Venezuela Humans
psANCOM.USA.Venezuela <- subset_samples(psH.prune, description =="Venezuela" | description =="USA")
psANCOM.USA.Venezuela.rare <- rarefy_even_depth(psANCOM.USA.Venezuela, rngseed=1, sample.size=0.9*min(sample_sums(psANCOM.USA.Venezuela)), replace=F)

```


## Start of analysis for adult USA Humans and Maya anoles
## Ordinate to a distance matrix -> plot via different methods -> run stats on the distance matrix -> make distance boxplots in qiime

## Maya anoles

``` {r, echo = T, eval = T}
# Distance matrices for Jacc and Bray
psM.rare.distance.bray <- phyloseq::distance(psM.rare, method = "bray", binary = FALSE)
psM.rare.distance.dice <- phyloseq::distance(psM.rare, method = "bray", binary = TRUE)

#convert phyloseq distance matrix to a tsv for qiime.

psUM.distance.bray.matrix <- as.matrix(psUM.distance.bray)
psUM.distance.bray.df  <- as.data.frame(psUM.distance.bray.matrix) %>% rownames_to_column(var = " ")
write_tsv(psUM.distance.jacc.df, file ="psUM_dist_jacc.tsv")

#ordinate the distance matrices using PCoA
psUM.PCoA.dice = ordinate(psUM.rare, "PCoA", distance= psUM.distance.dice)
psUM.PCoA.bray = ordinate(psUM.rare, "PCoA", distance= psUM.distance.bray)

#plot the ordination
psUM.PCoA.dice.plot <- plot_ordination(psUM.rare, psUM.PCoA.dice, color="description", title = "Dice")
psUM.PCoA.bray.plot <- plot_ordination(psUM.rare, psUM.PCoA.bray, color="description", title = "Bray")



psUM.NMDS.bray = ordinate(psUM.rare, "NMDS", distance= psUM.distance.bray)
psUM.NMDS.dice = ordinate(psUM.rare, "NMDS", distance= psUM.distance.dice)




ggsave("figures/psUM_dice_ordmethods.pdf", device = "pdf", width = 11, height = 8)
```

## USA/Maya
``` {r, echo = T, eval = T}
# Distance matrices for Jacc and Bray
psUM.distance.bray <- phyloseq::distance(psUM.rare, method = "bray", binary = FALSE)
psUM.distance.dice <- phyloseq::distance(psUM.rare, method = "bray", binary = TRUE)

#convert phyloseq distance matrix to a tsv for qiime.

psUM.distance.bray.matrix <- as.matrix(psUM.distance.bray)
psUM.distance.bray.df  <- as.data.frame(psUM.distance.bray.matrix) %>% rownames_to_column(var = " ")
write_tsv(psUM.distance.jacc.df, file ="psUM_dist_jacc.tsv")


#ordinate the distance matrices using PCoA
psUM.PCoA.dice = ordinate(psUM.rare, "PCoA", distance= psUM.distance.dice)
psUM.PCoA.bray = ordinate(psUM.rare, "PCoA", distance= psUM.distance.bray)

#plot the ordination
psUM.PCoA.dice.plot <- plot_ordination(psUM.rare, psUM.PCoA.dice, color="description", title = "Dice")
psUM.PCoA.bray.plot <- plot_ordination(psUM.rare, psUM.PCoA.bray, color="description", title = "Bray")


# Try some of the different ordination methods

psUM.NMDS.bray = ordinate(psUM.rare, "NMDS", distance= psUM.distance.bray)
psUM.NMDS.dice = ordinate(psUM.rare, "NMDS", distance= psUM.distance.dice)



psUM.NMDS.bray.plot = plot_ordination(psUM.rare, psUM.NMDS.bray, color="description")
psUM.NMDS.dice.plot = plot_ordination(psUM.rare, psUM.NMDS.dice, color="description")
psUM.NMDS.bray.plot + scale_color_manual(values = c("#08306B", "#2171B5", "#C6DBEF", "#6BAED6", "#e3c342"))



ggsave("figures/psUM_dice_ordmethods.pdf", device = "pdf", width = 11, height = 8)
```

## USA/Maya/Coyotes with duo
``` {r, echo = T, eval = T}
sample_data(psUMC.duo.rare)

# Distance matrices for Jacc and Bray
psUMC.duo.distance.dice <- phyloseq::distance(psUMC.duo.rare, method = "bray", binary= TRUE)
psUMC.duo.distance.bray <- phyloseq::distance(psUMC.duo.rare, method = "bray", binary= FALSE)


#ordinate the distance matrices using PCoA
psUMC.PCoA.jacc = ordinate(psUMC, "PCoA", distance= psUMC.distance.jacc)
psUMC.duo.PCoA.bray = ordinate(psUMC.duo.rare, "PCoA", distance= psUMC.duo.distance.dice)

#plot the ordination
psUMC.PCoA.jacc.plot <- plot_ordination(psUMC, psUMC.PCoA.jacc, color="description", title = "Jacc")
psUMC.duo.PCoA.bray.plot <- plot_ordination(psUMC.duo.rare, psUMC.duo.PCoA.bray, color="description", title = "Bray")




ggsave("figures/psUMC.duo_dice_ordmethods.pdf", device = "pdf", width = 11, height = 8)


#Qiime distance boxplots

psUMC.duo.distance.bray.matrix <- as.matrix(psUMC.duo.distance.bray)
psUMC.duo.distance.bray.df  <- as.data.frame(psUMC.duo.distance.bray.matrix) %>% rownames_to_column(var = " ")
write_tsv(psUMC.duo.distance.bray.df, file ="psUM_duo_dist_bray.tsv")
```


## USA/Maya/Coyotes
``` {r, echo = T, eval = T}
psUMC.rare

# Distance matrices for Jacc and Bray
psUMC.distance.dice <- phyloseq::distance(psUMC.rare, method = "bray", binary= TRUE)
psUMC.distance.bray <- phyloseq::distance(psUMC.rare, method = "bray", binary= FALSE)

#ordinate the distance matrices using NMDS
psUMC.NMDS.dice = ordinate(psUMC.rare, "NMDS", distance= psUMC.distance.dice)
psUMC.NMDS.bray = ordinate(psUMC.rare, "NMDS", distance= psUMC.distance.bray)

#plot the ordination
psUMC.NMDS.dice.plot <- plot_ordination(psUMC.rare, psUMC.NMDS.dice, color="description", title = "Dice")
psUMC.NMDS.bray.plot <- plot_ordination(psUMC.rare, psUMC.NMDS.bray, color="description", title = "Bray")


#test colors

#00f0c6 - light green 
#003b31 - dark green
#FFC107 - gold

psUMC.NMDS.dice.plot + scale_color_manual(values = c("#14548c", "#70bdff", "#1d86e0", "#0a365c", "#D81B60", "#851e44", "#FFC107")) + theme_classic() 

#psUMC.NMDS.dice.plot + scale_color_manual(values = c("#00856e", "#00f0c6", "#00bf9e", "#003b31", "#70bdff", "#14548c", "#FFC107")) + theme_classic() 


psUMC.NMDS.bray.plot + scale_color_manual(values = c("#14548c", "#70bdff", "#1d86e0", "#0a365c", "#D81B60", "#851e44", "#FFC107")) + theme_classic()

levels(psUMC.NMDS.dice.plot$description)


#Qiime distance boxplots

psUMC.distance.dice.matrix <- as.matrix(psUMC.distance.dice)
psUMC.distance.dice.df  <- as.data.frame(psUMC.distance.dice.matrix) %>% rownames_to_column(var = " ")
write_tsv(psUMC.distance.dice.df, file ="distance_boxplots/psUMC_dist_dice.tsv")

psUMC.distance.bray.matrix <- as.matrix(psUMC.distance.bray)
psUMC.distance.bray.df  <- as.data.frame(psUMC.distance.bray.matrix) %>% rownames_to_column(var = " ")
write_tsv(psUMC.distance.bray.df, file ="distance_boxplots/psUMC_dist_bray.tsv")



```




## Venez and Malawi distance boxplot inputs.

``` {r, echo = T, eval = T}
psVMC.rare

# Distance matrices for Jacc and Bray
psVMC.distance.dice <- phyloseq::distance(psVMC.rare, method = "bray", binary= TRUE)
psVMC.distance.bray <- phyloseq::distance(psVMC.rare, method = "bray", binary= FALSE)




#Qiime distance boxplots

psVMC.distance.dice.matrix <- as.matrix(psVMC.distance.dice)
psVMC.distance.dice.df  <- as.data.frame(psVMC.distance.dice.matrix) %>% rownames_to_column(var = " ")
write_tsv(psVMC.distance.dice.df, file ="distance_boxplots/psVMC_dist_dice.tsv")

psVMC.distance.bray.matrix <- as.matrix(psVMC.distance.bray)
psVMC.distance.bray.df  <- as.data.frame(psVMC.distance.bray.matrix) %>% rownames_to_column(var = " ")
write_tsv(psVMC.distance.bray.df, file ="distance_boxplots/psVMC_dist_bray.tsv")



psMAMC.rare

# Distance matrices for Jacc and Bray
psMAMC.distance.dice <- phyloseq::distance(psMAMC.rare, method = "bray", binary= TRUE)
psMAMC.distance.bray <- phyloseq::distance(psMAMC.rare, method = "bray", binary= FALSE)



#Qiime distance boxplots

psMAMC.distance.dice.matrix <- as.matrix(psMAMC.distance.dice)
psMAMC.distance.dice.df  <- as.data.frame(psMAMC.distance.dice.matrix) %>% rownames_to_column(var = " ")
write_tsv(psMAMC.distance.dice.df, file ="distance_boxplots/psMAMC_dist_dice.tsv")

psMAMC.distance.bray.matrix <- as.matrix(psMAMC.distance.bray)
psMAMC.distance.bray.df  <- as.data.frame(psMAMC.distance.bray.matrix) %>% rownames_to_column(var = " ")
write_tsv(psMAMC.distance.bray.df, file ="distance_boxplots/psMAMC_dist_bray.tsv")



```



## ANCOMII
``` {r, echo = T, eval = T}
library(nlme)
library(tidyverse)
library(ggplot2)
library(compositions)
source("scripts/ancom_v2.1.R")

testps <- subset_samples(psC.prune, description == "Peri-urban coyote fecal samples" | description =="Urban coyote fecal samples")
testps.rare <- rarefy_even_depth(testps, rngseed=1, sample.size=0.9*min(sample_sums(testps)), replace=F)

sample_data(testps)

testps
sample_data(testps)
otu_table(testps)
tax_table(testps)

# First need to transform the data into ancom form. just use as.data.frame. ANCOM wants the otu_data to have rownames and then it wants the rownames of the metadata converted to a column named sample.ID. If we want the rowname to column command to work you must first make the meta_data class a data.frame
otu_data = as.data.frame(otu_table(psANCOM.USA.Venezuela.rare))
meta_data = as.data.frame(sample_data(psANCOM.USA.Venezuela.rare))
class(meta_data) <- "data.frame"
meta_data = rownames_to_column(meta_data, var = "Sample.ID")
otu_data = otu_data[rowSums(otu_data[])>0,]
otu_id = rownames(otu_data)

# Step 1: Data preprocessing

feature_table = otu_data; sample_var = "Sample.ID"; group_var = NULL
out_cut = 0.05; zero_cut = 0.90; lib_cut = 1000; neg_lb = FALSE
prepro = feature_table_pre_process(feature_table, meta_data, sample_var, group_var, 
                                   out_cut, zero_cut, lib_cut, neg_lb)
feature_table = prepro$feature_table # Preprocessed feature table
meta_data = prepro$meta_data # Preprocessed metadata
struc_zero = prepro$structure_zeros # Structural zero info

# Step 2: ANCOM

main_var = "description"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = NULL
t_start = Sys.time()
res = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula)
t_end = Sys.time()
t_run = t_end - t_start # around 30s


# Step 3: Volcano Plot

# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1), 0.5 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6", "detected_0.5")

# Annotation data
dat_ann = data.frame(x = min(res$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")

fig = res$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE) + ggtitle("USA_Venez_ASV")
fig


#Filter data based on cutoff value

cut_off["detected_0.5"]

res
res$fig$data

fig_MU_MR_Genus
fig_MU_MR_Species
MUABBres$fig$data
fig_Coyotes_ASV

ggsave("figures/fig_USA_Malawi_ASV", device = "pdf")

fig_MU_MF_ASV
fig_MU_MR_ASV
fig_MU_ABB_ASV
fig_coyote_ASV
fig_USA_Venz_ASV
fig_USA_Malawi_ASV

coyoteres
MUABBres
MUMFres
MUMRres

filter(coyoteres$out, detected_0.6 == TRUE)
filter(MUABBres$out, detected_0.7 == TRUE)
filter(MUMFres$out, detected_0.7 == TRUE)
filter(MUMRres$out, detected_0.7 == TRUE)



# diff between MU and ABnB
as.data.frame(tax_table(psM.prune)) %>% rownames_to_column(var = "ASVID") %>% filter(ASVID == "ASV16414") #%>% write_csv(file ="figures/MU_ABnB_diffabundance.csv")
# diff between MU and MF
as.data.frame(tax_table(psM.prune)) %>% rownames_to_column(var = "ASVID") %>% filter(ASVID == "ASV6593" | ASVID == "ASV7025" | ASVID == "ASV11740") #%>% write_csv(file ="figures/MU_MF_diffabundance.csv")
# diff between MU and MR
as.data.frame(tax_table(psM.prune)) %>% rownames_to_column(var = "ASVID") %>% filter(ASVID %in% c("ASV4663", "ASV6593", "ASV7025", "ASV7280", "ASV8582", "ASV13425", "ASV16638", "ASV22190", "ASV26624", "ASV26736")) #%>% write_csv(file ="figures/MU_MR_diffabundance.csv")
# diff between coyotes
as.data.frame(tax_table(psC.prune)) %>% rownames_to_column(var = "ASVID") %>% filter(ASVID == "ASV19872" | ASVID == "ASV28851") #%>% write_csv(file ="figures/Coyotes_diffabundance.csv")

```




## Try to streamline these boxplots a bit with a function to do the data cleaning
``` {r, echo = T, eval = T}



# Functiont to hopefully make boxplots for me. Resfile is just whatever the res output is named, Wcutoff should be in the "detected_0.X" format ( I guess you could also save it to a variable and not use the ""), functionlevels should be a matrix with the order of sample names you want on the x axis, outname is desired boxplot name.

getASVboxplots <- function(resfile, Wcutoff, functionlevels, outname) {
  
  
  function.matrix <- filter(resfile$out, Wcutoff == TRUE) %>% select(taxa_id) %>% as.matrix()
  function.matrix <- as.character(c(function.matrix))
  function_ps <- prune_taxa(c(function.matrix), psAll.rare)
  functionmeta <- as.data.frame(sample_data(function_ps))
  class(functionmeta) <- "data.frame"
  functionmeta <- rownames_to_column(functionmeta, var = "sample")
  functionabundance <- as.data.frame(otu_table(function_ps)) %>%  t() %>% as.data.frame() %>% rownames_to_column(var = "sample")
  function_boxplot <- dplyr::left_join(functionabundance, functionmeta, by = "sample")
  function_boxplot <- pivot_longer(function_boxplot, function.matrix, names_to = "ASV", values_to = "abundance")
  function_boxplot$description <- as.character(function_boxplot$description)
  function_boxplot$description <- factor(function_boxplot$description, levels = c(functionlevels))
  
  outname <-  function_boxplot %>% ggplot(aes(x=description, y=abundance)) + geom_boxplot(aes(fill = description), width = .5) + theme_classic() + theme(axis.text.x = element_text(angle = -45, vjust = .75, hjust = 0)) + facet_wrap(~function_boxplot$ASV, scales = "free_y") + scale_fill_manual(values = group.colors)
  
  outname
}

anolelevels = c("Mayaguez Urban","AirBnB","Mayaguez Forest","Maricao Rocks","Urban coyote fecal samples", "Peri-urban coyote fecal samples", "USA", "Venezuela", "Malawi")
group.colors <- c("Maricao Rocks" = "#ffd3ff", "Mayaguez Forest" = "#ff86ff", "AirBnB" = "#ab00ff", "Mayaguez Urban" = "#002cc0", "Peri-urban coyote fecal samples" = "#ff1e44", "Urban coyote fecal samples" = "#700000", "USA" = "#ffc107", "Venezuela" = "#c99906", "Malawi" = "#997405")

coyoteres
MUABBres
MUMFres
MUMRres

getASVboxplots(coyoteres, coyoteres$out$detected_0.6, anolelevels, testmyfunction)

ggsave("figures/meta_boxplots/coyoteres_ASVabundances.pdf", device = "pdf", width = 16, height = 11)

```



## ASV boxplots
``` {r, echo = T, eval = T}

ASV6593
psAll.rare

ASV6593_ps <- prune_taxa("ASV6593", psAll.rare)

tax_table(ASV6593_ps)

ASV6593meta <- as.data.frame(sample_data(ASV6593_ps))
class(ASV6593meta) <- "data.frame"
ASV6593meta <- rownames_to_column(ASV6593meta, var = "sample")


ASV6593abundance <- as.data.frame(otu_table(ASV6593_ps)) %>% gather(key = "sample", value = "abundance")
otu_table(ASV6593_ps)
ASV6593meta
ASV6593abundance

ASV6593_boxplot <- dplyr::left_join(ASV6593abundance, ASV6593meta, by = "sample")
ASV6593_boxplot$description <- as.character(ASV6593_boxplot$description)
ASV6593_boxplot

ASV6593_boxplot$description <- factor(ASV6593_boxplot$description,levels = c("Maricao Rocks", "Mayaguez Forest", "AirBnB", "Mayaguez Urban", "Peri-urban coyote fecal samples", "Urban coyote fecal samples", "USA", "Venezuela", "Malawi"))

ASV6593_boxplot %>% ggplot(aes(x=description, y=abundance)) + geom_boxplot(fill = c(group.colors), width = .5) + theme_classic() + theme(axis.text.x = element_text(angle = -45, vjust = .75, hjust = 0))




ggsave("figures/ASV6593_abundance.pdf", device = "pdf", width = 18, height = 8)


everything.levels <- c("Maricao Rocks", "Mayaguez Forest", "AirBnB", "Mayaguez Urban", "Peri-urban coyote fecal samples", "Urban coyote fecal samples", "USA", "Venezuela", "Malawi")
evenything.colors <- c("#ffd3ff", "#ff86ff", "#ab00ff", "#002cc0", "#ff1e44", "#700000", "#ffc107", "#c99906", "#997405")

# have the opject now need to do math and then plot
#count(ASV6593_boxplot, vars = "description")
#sum(ASV6593_boxplot$abundance)
#ASV6593_boxplot
#test <- filter(ASV6593_boxplot, description == "AirBnB")
#sum(test$abundance)

```



## Distance boxplot figures

``` {r, echo = T, eval = T}
# Get all the boxplot data and make it into something usable
all_boxplots =  read_tsv("distance_boxplots/VMC/VMC_dice_Distances.txt", col_names = FALSE)
all_boxplots <- as.data.frame(t(all_boxplots))
colnames(all_boxplots) <- all_boxplots[1,]
all_boxplots <- all_boxplots[-1,]
all_boxplots <-remove_rownames(all_boxplots)
all_boxplots


# subset what we want and then make it graphable
figure_boxplots <- select(all_boxplots, "Mayaguez Urban" = Venezuela_vs._Mayaguez_Urban, "AirBnB" = AirBnB_vs._Venezuela, "Mayaguez Forest" = Mayaguez_Forest_vs._Venezuela, "Maricao Rocks" = Maricao_Rocks_vs._Venezuela, "Urban coyote fecal" = Urban_coyote_fecal_samples_vs._Venezuela, "Peri-urban coyote fecal" = "Venezuela_vs._Peri-urban_coyote_fecal_samples")
figure_boxplots <- figure_boxplots %>% pivot_longer(everything(), names_to = "type", values_to = "distance")
figure_boxplots$distance <- as.numeric(figure_boxplots$distance)
figure_boxplots <- figure_boxplots %>% drop_na()
figure_boxplots

figure_boxplots$type <- factor(figure_boxplots$type,levels = c("Mayaguez Urban", "AirBnB", "Mayaguez Forest", "Maricao Rocks", "Urban coyote fecal", "Peri-urban coyote fecal"))

ggplot(figure_boxplots) + geom_boxplot(aes(x = type, y = distance)) + ggtitle("Dice Distance from Venezuela Human Adults")

ggsave("figures/Dice_Venezuela_distance_boxplots.pdf", device = "pdf", width = 14, height = 9)
```



## presence adsence of specific ASVs between urban/humans and natural/humans

``` {r, echo = T, eval = T}

psAll.rare

otu_table(psAll.rare)

psAll.rare.meta <- as.data.frame(sample_data(psAll.rare))
class(psAll.rare.meta) <- "data.frame"
psAll.rare.meta <- rownames_to_column(psAll.rare.meta, var = "sample")
psAll.rare.abundance <- as.data.frame(otu_table(psAll.rare)) %>%  t() %>% as.data.frame() %>% rownames_to_column(var = "sample")

psAll.rare.abundance <- dplyr::left_join(psAll.rare.abundance, psAll.rare.meta, by = "sample")
psAll.rare.abundance <- pivot_longer(psAll.rare.abundance, rownames(otu_table(psAll.rare)), names_to = "ASV", values_to = "abundance")


psAll.rare.abundance %>% group_by(description) %>% summarize(totalabun = sum(abundance))


psAll.grouped <- psAll.rare.abundance %>% select(description,rownames(otu_table(psAll.rare)))
psAll.grouped

psAll.grouped <- aggregate(. ~ description, data=psAll.grouped, FUN=sum)
psAll.grouped <- pivot_longer(psAll.grouped, rownames(otu_table(psAll.rare)), names_to = "ASV", values_to = "abundance")
psAll.grouped.filter <- psAll.grouped %>% filter(description == "Maricao Rocks" | description == "Mayaguez Forest" | description == "AirBnB" | description == "Mayaguez Urban" | description == "USA")

psAll.grouped.coyote <- psAll.grouped %>% filter(description == "Peri-urban coyote fecal samples" | description == "Urban coyote fecal samples" | description == "USA")
psAll.grouped.coyote
psAll.grouped.filter %>% filter(abundance != 0 & description == c("Mayaguez Urban", "USA", "AirBnbB"))
psAll.grouped.filter %>% filter(abundance == 0 & description == c("Mayaguez Forest", "Maricao Rocks"))


test <- psAll.grouped.filter %>% filter(abundance == 0 & description == "Mayaguez Urban") %>% select(-description, -abundance)
test2 <- psAll.grouped.filter %>% filter(abundance != 0 & description == "USA") %>% select(-description, -abundance)
test3 <- psAll.grouped.filter %>% filter(abundance != 0 & description == "AirBnB") %>% select(-description, -abundance)
present.match <- match_df(test, test3) #%>% match_df(test3)


neg.test <- psAll.grouped.filter %>% filter(abundance == 0 & description == "Mayaguez Forest") %>% select(-description, -abundance)
neg.test2 <- psAll.grouped.filter %>% filter(abundance != 0 & description == "Maricao Rocks") %>% select(-description, -abundance)

neg.match <- match_df(neg.test, neg.test2) %>% match_df(test2)


match_df(present.match, neg.match)

#This works without having to do the other matching
match_df(test, test2) %>% match_df(test3) %>% match_df(neg.test) %>% match_df(neg.test2)


coyote.urban.match <- psAll.grouped.coyote %>% filter(abundance == 0 & description == "Urban coyote fecal samples") %>% select(-description, -abundance)
coyote.peri.match <- psAll.grouped.coyote %>% filter(abundance != 0 & description == "Peri-urban coyote fecal samples") %>% select(-description, -abundance)

coyote.urban.match


match_df(test, test2) %>% match_df(neg.test2) %>% write_csv("figures/tables/test")



asvlist <- read.csv("figures/tables/asvlist.csv")
asvlist <- as.matrix(asvlist)
asvlist
asv.list.seqs <- refseq(psAll.rare)

asv.list.seqs<- as.data.frame(asv.list.seqs)

asv.list.seqs <- rownames_to_column(asv.list.seqs, var = "ASV")

filter(asv.list.seqs, ASV %in% asvlist) %>% write_csv("list_of_asv.csv")


```


``` {r, echo = T, eval = T}
psUMC.rare
testasvs <- psotu2veg(psUMC.rare)
testmeta <-  pssd2veg(psUMC.rare)
# Distance matrices for Jacc and Bray
psUMC.distance.dice
psUMC.distance.bray

testmeta
M.vegdist=vegdist(testasvs, distance="bray")


adonis(psUMC.distance.dice ~ description, data = testmeta, permutations = 1000)
betadisper(psUMC.distance.dice, testmeta$description)
anosim(psUMC.distance.dice, testmeta$description, permutations = 1000)
```


















