---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r, echo = T, eval = T}

#load packages

library(dada2)
library(phyloseq)
library(Biostrings)
library(ggplot2)
library(dbplyr)
library(tidyverse)
library(qiime2R)
library(plyr)
library(ANCOMBC)
library(microbiome)
library(rmeta)
library(readr)
library(vegan)
library(biomformat)
library(nlme)
library(compositions)
source("scripts/ancom_v2.1.R")

setwd("/workdir/bd429/R/Anole_paper")
```

## Make functions to convert Phyloseq objects into vegan objects

``` {r, echo = T, eval = T}
# convert the sample_data() within a phyloseq object into a vegan compatible data object
pssd2veg <- function(physeq) {
  sd <- sample_data(physeq)
  return(as(sd,"data.frame"))
}

#convert the asv table (otu_table) within a phyloseq object into a vegan compatible data object
psotu2veg <- function(physeq) {
  OTU <- otu_table(physeq)
  if (taxa_are_rows(OTU)) {
    OTU <- t(OTU)
  }
  return(as(OTU, "matrix"))
}


options(max.print = 5000)
```


## qiime 2 R import and conversion into phyloseq object
``` {r, echo = T, eval = T}

#import qiime2 data into a phyloseq object
allps<-qza_to_phyloseq(features="qiime2/filter_merged_feature-table.qza", taxonomy = "qiime2/mergedtaxonomy.qza", metadata="qiime2/qiime2_meta_filtered1.txt")

# convert sequences to ASV names and create new field to keep the reference sequences
dna <- Biostrings::DNAStringSet(taxa_names(allps))
names(dna) <- taxa_names(allps)
allps <- merge_phyloseq(allps, dna)
taxa_names(allps) <- paste0("ASV", seq(ntaxa(allps)))


allps <- prune_samples(sample_sums(allps)>=10000, allps)
allps <- subset_samples(allps, original_ID !="AF107" & original_ID !="AF008" & original_ID !="AF086" & original_ID !="53F")



```



## split all datasets into own subsets and filter like children and non cristatellus
``` {r, echo = T, eval = T}

psA <- subset_samples(allps, env_feature == "AnoleFeces") %>% subset_samples(description == "Mayaguez Forest" | description == "Mayaguez Urban" | description =="Maricao Rocks" | description =="AirBnB") %>% subset_samples(scientific_name != "gundlachi")


psC <- subset_samples(allps, study == "coyote") 


psY <- subset_samples(allps, study == "yatsunenko") %>% subset_samples(life_stage != "child")

psH <- subset_samples(allps, study == "himalaya")

psT <- subset_samples(allps, study == "tibet")

psM <- subset_samples(allps, study == "cameroon")

psS <- subset_samples(allps, study == "sparrow")

psanimals.prune <- merge_phyloseq(psC, psA, psS)
animals.ancom <- subset_samples(psanimals.prune, ANCOM_random != "None" & scientific_name != "Homo sapien")


psall.prune <- merge_phyloseq(psA, psC, psY, psH, psT, psM, psS)
psall.prune.rare = rarefy_even_depth(psall.prune, rngseed=1, sample.size=10000, replace=F)

#psall.rare.distance.bray <- phyloseq::distance(psall.prune.rare, method = "bray", binary = FALSE)
#psall.NMDS.dice <- ordinate(psall.prune.rare, "NMDS", distance= psall.rare.distance.bray)
#plot_ordination(psall.prune.rare, psall.NMDS.dice, color="LifestyleSpecies", title = "nmds")
#ggsave("figures/revisions/pruned_Bray_NMDS.pdf", device = "pdf", width = 13, height = 11)


ancom_refseq <- as.data.frame(refseq(allps)) %>% rownames_to_column(var = "taxa_id") %>% mutate(sequence = x) %>% select(taxa_id, sequence)
Bact_table <- as.data.frame(tax_table(allps)) %>% rownames_to_column(var = "taxa_id")
Bact_table <- left_join(Bact_table, ancom_refseq)
```



``` {r, echo = T, eval = T}

# add unrarefied physloseq object
ANCOMps <- psA

# First need to transform the data into ancom form. Use as.data.frame. ANCOM wants the otu_data to have rownames and then it wants the rownames of the metadata converted to a column named sample.ID. If we want the rowname to column command to work you must first make the meta_data class a data.frame
otu_data = as.data.frame(otu_table(ANCOMps))
meta_data = as.data.frame(sample_data(ANCOMps))
class(meta_data) <- "data.frame"
meta_data = rownames_to_column(meta_data, var = "Sample.ID")
otu_data = otu_data[rowSums(otu_data[])>0,]
otu_id = rownames(otu_data)


# Step 1: Data preprocessing

feature_table = otu_data; sample_var = "Sample.ID"; group_var = NULL
out_cut = 0.05; zero_cut = 0.90; lib_cut = 1000; neg_lb = FALSE
prepro = feature_table_pre_process(feature_table, meta_data, sample_var, group_var, 
                                   out_cut, zero_cut, lib_cut, neg_lb)
feature_table = prepro$feature_table # Preprocessed feature table
meta_data = prepro$meta_data # Preprocessed metadata
struc_zero = prepro$structure_zeros # Structural zero info

# Step 2: ANCOM

main_var = "ANCOM_random"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = NULL ; lme_control = NULL
res = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula, lme_control)



# Step 3: Volcano Plot

# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1), 0.5 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6", "detected_0.5")

# Annotation data + figure
dat_ann = data.frame(x = min(res$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")

ANCOMfig = res$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
ANCOMfig


significant_5 <- res$fig$data %>% mutate(detected_0.5 = ifelse(y >= cut_off[5], "TRUE", "FALSE")) %>% mutate(detected_0.7 = ifelse(y >= cut_off[3], "TRUE", "FALSE")) %>% mutate(W = y) %>% filter(detected_0.5 == TRUE)
rownames(significant_5) <- NULL
significant_5

ancom_table <- merge(x = significant_5, y = Bact_table, by = "taxa_id") %>% as.data.frame() %>% mutate(clr_mean_diff = x) %>% mutate(description = ifelse(clr_mean_diff > 0, "Urban", "Rural")) %>% mutate(study = "Animals") %>% select(taxa_id, W, clr_mean_diff, detected_0.5, detected_0.7, description, Kingdom, Phylum, Order, Family, Genus, study, sequence)

#Anole_ancom_fig <- ANCOMfig
#AnoleRes <- res
#anole_ancom <- ancom_table


sparrowRes
sparrow_ancom_fig
sparrow_ancom

tibet_tradurbanRes
tibet_tradurban_ancom_fig
tibet_tradurban_ancom

tibet_semiurbanRes
tibet_semiurban_ancom_fig
tibet_semiurban_ancom

cameroon_ruralRes
cameroon_rural_ancom_fig
cameroon_rural_ancom

cameroon_semiurbanRes
cameroon_semiurban_ancom_fig
cameroon_semi_ancom

AnoleRes
Anole_ancom_fig
anole_ancom

venezuelaRes
venezuela_ancom_fig
venezuela_ancom

malawiRes
malawi_ancom_fig
malawi_ancom

coyoteRes
coyote_ancom_fig
coyote_ancom


ANCOM_allresults <- rbind(sparrow_ancom, tibet_tradurban_ancom, tibet_semiurban_ancom, cameroon_rural_ancom, cameroon_semi_ancom, anole_ancom, venezuela_ancom, malawi_ancom, coyote_ancom)

write_csv(allanimal_ancom, file = "allanimal_ancom.csv")

```

## ANCOM of all animals with covariable. Needs less restrictive memory requirements, this code was run on the command line with the commands "ulimit -s 65535" and "R --max-ppsize=500000".
```{r}

ANCOMps <- animals.ancom


#save R data so that it can be called through the command line
save(ANCOMps, file = "animals_ANCOM.R")
# First need to transform the data into ancom form. Pretty easy just use as.data.frame. ANCOM wants the otu_data to have rownames and then it wants the rownames of the metadata converted to a column named sample.ID. If we want the rowname to column command to work you must first make the meta_data class a data.frame

#Starting here ran on command line with less restrictive memory limits
load("animals_ANCOM.R")
otu_data = as.data.frame(otu_table(ANCOMps))
meta_data = as.data.frame(sample_data(ANCOMps))
class(meta_data) <- "data.frame"
meta_data = rownames_to_column(meta_data, var = "Sample.ID")
otu_data = otu_data[rowSums(otu_data[])>0,]
otu_id = rownames(otu_data)


# Step 1: Data preprocessing

feature_table = otu_data; sample_var = "Sample.ID"; group_var = NULL
out_cut = 0.05; zero_cut = 0.90; lib_cut = 1000; neg_lb = FALSE
prepro = feature_table_pre_process(feature_table, meta_data, sample_var, group_var, 
                                   out_cut, zero_cut, lib_cut, neg_lb)
feature_table = prepro$feature_table # Preprocessed feature table
meta_data = prepro$meta_data # Preprocessed metadata
struc_zero = prepro$structure_zeros # Structural zero info

# Step 2: ANCOM
#ANCOM_random is the Urban or Rural designation for a sample in the metadata, it is not random.
main_var = "ANCOM_random"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL; rand_formula = "~1|scientific_name" ; lme_control = list(maxIter = 100, msMaxIter = 100, opt = "optim")
res2 = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method, 
            alpha, adj_formula, rand_formula, control = lme_control)


# Step 3: Volcano Plot

# Number of taxa except structural zeros
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1), 0.5 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6", "detected_0.5")

# Annotation data + figure
dat_ann = data.frame(x = min(res2$fig$data$x), y = cut_off["detected_0.7"], label = "W[0.7]")

ANCOMfig = res2$fig +  
  geom_hline(yintercept = cut_off["detected_0.7"], linetype = "dashed") + 
  geom_text(data = dat_ann, aes(x = x, y = y, label = label), 
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
ANCOMfig


#Make the supp table 
significant_5 <- res2$fig$data %>% mutate(detected_0.5 = ifelse(y >= cut_off[5], "TRUE", "FALSE")) %>% mutate(detected_0.7 = ifelse(y >= cut_off[3], "TRUE", "FALSE")) %>% mutate(W = y) %>% filter(detected_0.5 == TRUE)
rownames(significant_5) <- NULL


allanimal_ancom <- merge(x = significant_5, y = Bact_table, by = "taxa_id") %>% as.data.frame() %>% mutate(clr_mean_diff = x) %>% mutate(description = ifelse(clr_mean_diff > 0, "Urban", "Rural")) %>% mutate(study = "Animals") %>% select(taxa_id, W, clr_mean_diff, detected_0.5, detected_0.7, description, Kingdom, Phylum, Order, Family, Genus, study, sequence)


load("ancom_res.R")
load("ancom_fig.R")



#comboy <- otu_table(animals.ancom) %>% as.data.frame() %>% rownames_to_column(var = "taxa_id") %>% filter(taxa_id == "ASV17782") %>% as.data.frame() %>% pivot_longer(!taxa_id, names_to = "id", values_to = "count")

#combo <- as(phyloseq::sample_data(animals.ancom), "data.frame") %>% rownames_to_column(var = "id") %>% select(id, study, ANCOM_random) %>% left_join(., comboy)

#write_csv(combo, file = "combopos.csv")
```



## presence absence of ASVs + random subsampling for the studies that need it.
```{r}


#Randomly create lists of samples 
sampling <- sample_data(psall.prune.present) %>% `class<-`("data.frame") %>% rownames_to_column(var = "sample") %>% filter(LifestyleSpecies == "UrbanSparrow") %>% select(original_ID) %>% as.matrix() %>% sample(23)

subsample_data <- subset_samples(psall.prune.present, original_ID %in% sampling | LifestyleSpecies == "RuralSparrow" | LifestyleSpecies == "RuralHuman" | LifestyleSpecies == "UrbanHuman")



otuframe <- t(otu_table(subsample_data)) %>% as.data.frame() %>% rownames_to_column(var = "sample")
sample_meta_frame <- as.data.frame(sample_data(subsample_data)) %>% `class<-`("data.frame") %>% rownames_to_column(var = "sample") %>% select(sample, original_ID, LifestyleSpecies, study)
#below command is to homogenize the names of the various groups
sample_meta_frame$LifestyleSpecies[sample_meta_frame$LifestyleSpecies == "RuralHuman" | sample_meta_frame$LifestyleSpecies == "UrbanHuman"] <- "Human"

merge_meta_otu <- merge(x = otuframe, y = sample_meta_frame, by = "sample", all.x = TRUE)
merge_meta_otu <- merge_meta_otu %>% select(LifestyleSpecies, rownames(otu_table(subsample_data)))

aggregate_OTUsums <- aggregate(. ~ LifestyleSpecies, data=merge_meta_otu, FUN=sum)
aggregate_OTUsums <- pivot_longer(aggregate_OTUsums, rownames(otu_table(subsample_data)), names_to = "ASV", values_to = "abundance")


present1 <- aggregate_OTUsums %>% filter(abundance != 0 & LifestyleSpecies == "RuralSparrow") %>% select(-LifestyleSpecies, -abundance)
present2 <- aggregate_OTUsums %>% filter(abundance != 0 & LifestyleSpecies == "Human") %>% select(-LifestyleSpecies, -abundance)
absent <- aggregate_OTUsums %>% filter(abundance == 0 & LifestyleSpecies == "UrbanSparrow") %>% select(-LifestyleSpecies, -abundance)


match_df(present1, present2) %>% match_df(absent) %>% write_csv("figures/revisions/RuralSparrow_Human_UrbanSparrow_R10.csv")



#combined all the pres/absent files in exel then read back into R to append the ASV sequences.
read.csv("figures/revisions/Presence_absence.csv", header = TRUE) %>% mutate(taxa_id = ASV) %>% merge(x = ., y = ancom_refseq, by = "taxa_id", all.x = TRUE) %>% select(Catergories, ASV, Replicate, sequence) %>% write.csv(file = "figures/revisions/Presence_absence_seqs.csv")

```

















